Marlowe: implementing and analysing financial contracts on blockchain

### 2019-12-09 papers/marlowe2019/paper.tex
New sections,

Because Marlowe is a DSL, we are able to build special purpose tools and techniques to support it. Crucially, in a financial environment we are able to \emph{exhaustively analyse contracts} without executing them, so that we can, for instance, check whether any particular contract is able to make all the payments it should: in the case it is not, we get an explicit example of how it can fail. This analysis, reported in Section \ref{sec:analysis}, is built into the Marlowe Playground. This section also covers our use of formal verification to prove properties of the implementation of Marlowe, including a guarantee that ``money in $=$ money out'' for all contracts. 
% PLS: Should we divide the analysis/formalisation section in two?

@@

The Marlowe interpreter first validates the current contract state: i.e.\
we check that the contract locks at least what it should according
to contract balances (the \emph{accounts} field in \emph{State}),
and all balances are strictly positive.
\footnote{
Using the Isabelle proof assistant, we have formally verified that given a state with positive balances,
it is impossible for any possible contract and inputs to result in non-positive balances.
This is described in more detail in Section \ref{sec:analysis}.}

@@

\section{Analysing Marlowe\label{sec:analysis}}

In this section, we give an overview of some of the work we have carried out with the aim of helping users be more confident that the Marlowe contracts they write do what they expect them to do.

We look at two complementary approaches: static analysis of concrete contracts (Section~\ref{subsec:static}) and formal verification of general properties about Marlowe (Section~\ref{subsec:verification}).

\subsection{Static analysis of contracts\label{subsec:static}}

@@

\subsection{Formal verification of semantics\label{subsec:verification}}

@@

\begin{itemize}
    \item It reduces the danger of Denegation of Service (DoS) attacks: because the number of valid inputs is limited, an attacker partcipant cannot arbitrarily block the contract by issuing an unbounded amount of useless \texttt{Transaction}s.
    \item It bounds the length of traces (the number of transactions) that symbolic execution (see Section~\ref{subsec:static}) needs to explore (more on this below).
\end{itemize}

>>>>

Because Marlowe is a DSL, we are able to build special purpose tools and techniques to support it. Crucially, in a financial environment we are able to \emph{exhaustively analyse contracts} without executing them, so that we can, for instance, check whether any particular contract is able to make all the payments it should: in the case it is not, we get an explicit example of how it can fail. This analysis, reported in Section \ref{sec:static}, is built into the Marlowe Playground. We are also able to use formal verification to prove properties of the implementation of Marlowe, including a guarantee that ``money in $=$ money out'' for all contracts; this is covered in Section~\ref{sec:verification}. 
% PLS: Should we divide the analysis/formalisation section in two? SJT: done

@@

The Marlowe interpreter first validates the current contract state: i.e.\
we check that the contract locks at least what it should according
to contract balances (the \emph{accounts} field in \emph{State}),
and all balances are strictly positive.
\footnote{
Using the Isabelle proof assistant, we have formally verified that given a state with positive balances,
it is impossible for any possible contract and inputs to result in non-positive balances.
This is described in more detail in Section \ref{sec:static}.}

@@

\section{Static analysis of contracts\label{sec:static}}

In the next two sections we give an overview of some of the work we have carried out with the aim of helping users be more confident that the Marlowe contracts they write do what they expect them to do.
We look at two complementary approaches: static analysis of concrete contracts in this section,  and formal verification of general properties about Marlowe in the next.

@@

\section{Formal verification of semantics\label{sec:verification}}

@@

\begin{itemize}
    \item It reduces the danger of Denegation of Service (DoS) attacks: because the number of valid inputs is limited, an attacker partcipant cannot arbitrarily block the contract by issuing an unbounded amount of useless \texttt{Transaction}s.
    \item It bounds the length of traces (the number of transactions) that symbolic execution (see Section~\ref{sec:static}) needs to explore (more on this below).
\end{itemize}
